\section{Basic Population Dynamics}

\paragraph{Difference between Population Dynamics and other models}

\paragraph{Types of models}

\paragraph{Single Species Model}~{\\
$t$ be time, changing with intervals $\tau$ (e.g. 1 year).\\
$N(t)$ be total population of the species at time $t$.\\
$\tilde{B}(t, \tau)$ be number of births in the population between $t$ and $t+\tau$, fertility/recruitment term.\\
$\tilde{D}(t, \tau)$ be number of deaths in the population between $t$ and $t+\tau$, mortality term, $\tilde{D} >= 0$.\\
$\tilde{b}(t)=\lim _{\tau \rightarrow 0} \tilde{B}(t, \tau) / \tau$ be numbers of births per unit time.\\
$\tilde{d}(t)=\lim _{\tau \rightarrow 0} \tilde{D}(t, \tau) / \tau$ be numbers of deaths per unit time.
}

\subparagraph{Discrete Time}~{
\begin{equation}
    N(t+\tau)=N(t)+\tilde{B}(t, \tau)-\tilde{D}(t, \tau)\label{eq1.1}
\end{equation}
}

\subparagraph{Continuous Time}~{when $\tau \rightarrow 0$, then
\begin{equation}
    \frac{\mathrm{d} N}{\mathrm{~d} t}=\lim _{\tau \rightarrow 0} \frac{N(t+\tau)-N(t)}{\tau}=\tilde{b}(t)-\tilde{d}(t)\label{eq1.2}
\end{equation}
}

\newpage
\subsection{Single Species Model: Malthusian Model}
\begin{assumption}[Malthusian Assumption]
    The acts of death and birth of animals (vegetation, cell) happen statistically independently of each other and independently of the current population size.
\end{assumption}


\subsubsection{Discrete Time Equation $N(t+\tau)=N(t)+\tilde{B}(t, \tau)-\tilde{D}(t, \tau)$}
Total number of births and deaths happening during timestep $\tau$ proportional to current population size, we have
$$
    \tilde{D}(t, \tau)=D(\tau) N(t), \quad \tilde{B}(t, \tau)=B(\tau) N(t)
$$
then \myref{eq1.1} becomes
\begin{align}
    N(t+\tau) & =(1+B(\tau)+D(\tau))N(t)\notag \\
    N(t+\tau) & =R(\tau)N(t)\label{eq1.3}
\end{align}
where
$$
    R(\tau) = 1+B(\tau)+D(\tau) = \frac{N(t+\tau)}{N(t)}
$$
is called \textbf{\emph{reproduction coefficient}}. $R(\tau)$ is size ratio and only $R(\tau)\geq 0$ make \textbf{biological sense}.$R(\tau) = 0$ is the trivial case: the population dies out within one time step.

\paragraph{Discrete Time Solution}~{}\\
to solve \myref{eq1.3}, we set $N(\tau)=R(\tau)N(0), N(0)\geq 0$ at $t=0$. And for all $t=n\tau$, we have
\begin{align}
     & t=0,N(\tau)=RN(0)\notag                              \\
     & t=\tau, N(2\tau)=R N(\tau)=R^{2} N(0)\notag          \\
     & t=2 \tau, N(3 \tau)=R N(2 \tau)=R^{3} N(0)\notag     \\
     & \qquad \qquad \qquad \quad...\notag                  \\
     & t=n \tau, N(n\tau)=RN(n\tau)=R^{n} N(0)\label{eq1.4}
\end{align}
That is, the population grows (for $R(\tau) > 1$) or decays (for $R(\tau) < 1$) in \emph{\textbf{Geometric Progression}}.

\newpage

\subsubsection{Continuous Time Equation $\frac{\mathrm{d} N}{\mathrm{~d} t}=\tilde{b}(t)-\tilde{d}(t)$}
Now we have the number of births and deaths happening continuously per unit time proportional to current population size.
$$
    \tilde{d}(t)=d N(t), \quad \tilde{b}(t)=b N(t)
$$
then \myref{eq1.2} becomes
\begin{align}
    \frac{\mathrm{d} N}{\mathrm{~d} t} & =\tilde{b}(t)-\tilde{d}(t)\notag \\
    \frac{\mathrm{d} N}{\mathrm{~d} t} & =rN(t) \label{eq1.5}
\end{align}
where
$$
    r=b-d
$$
is called \emph{\textbf{reproduction rate}}, r can be positive or negative.

\paragraph{Continuous Time Solution}~{}\\
We solve \myref{eq1.5} by \textbf{separation of variables}.
\begin{align}
    \frac{\mathrm{d}N}{N(t)}      & =r\mathrm{d}t \notag \\
    \int \frac{\mathrm{d}N}{N(t)} & =\int r d t \notag   \\
    \ln N(t)                      & =rt+C\notag          \\
    N(t)                          & = N(0)e^{rt}
\end{align}
where
$$
    N(t=0)=N(0)\geq 0
$$
This is, the population grows (for $r>0$) or decays (for $r<0$) \emph{\textbf{exponentially}}.

\newpage

\subsubsection{Link of Two Equation}

\paragraph{Discrete $\rightarrow$ Continuous}~{
$
    N(n\tau) = R^n(\tau) N(0)  \rightarrow N(t) = e^{rt}N(0)
$}\\
For small $\tau$,
\begin{align}
    \lim _{\tau \rightarrow 0} \frac{R(\tau)-1}{\tau} & = \lim _{\tau \rightarrow 0}\frac{B(\tau)-D(\tau)}{\tau}\notag \\
    R(\tau)                                           & \approx 1+\tau r\label{eq1.6}
\end{align}
For fixed time t,
$$
    n=\frac{t}{\tau} \rightarrow \infty
$$
Then we have
\begin{align*}
    N(n \tau) & =\lim_{\tau \rightarrow 0}R(\tau)^nN(0)                      \\
              & \approx \lim _{\tau \rightarrow 0}(1+r \tau)^{t / \tau} N(0) \\
              & =e^{rt}N(0)
\end{align*}


\paragraph{Continuous $\rightarrow$ Discrete}~{
$
    N(t) = e^{rt}N(0) \rightarrow N(n\tau) = R^n(\tau) N(0)
$
}\\
Let $t=n\tau$, we have
\begin{align*}
    N(t) & =N(n\tau)                \\
         & =e^{(r\tau)n}N(0)        \\
         & \approx (1+r\tau)^n N(0) \\
         & \approx R^n(\tau)N(0)
\end{align*}
by \myref{eq1.6} as $\tau \rightarrow 0$ and Taylor series.


\newpage
\subsection{Intraspecific Competition Model}

\emph{Competition}: increase of population size suppresses its reproduction.

\begin{assumption}[Intraspecific Competition Assumption]
    Reproduction per capita per unit of time is \textbf{NOT} constant. Instead, reproduction is depends on the current population size.
\end{assumption}

\subparagraph{Discrete Time}~{
    $$
    N_{t+1} = R(N_t)N_t
    $$
    \par{Competition when}
    $$
    dR/dN <0
    $$
}
\subparagraph{Continuous Time}~{
    $$
    dN/dt = r(N)N
    $$
    \par{Competition when}
    $$
    dr/dN <0
    $$
}

\subsubsection{Verhulst (Logistic) Model}
\paragraph{Equation}~{}\\
A simplest guess of $r(N)$ by Verhulst is 
$$
r=r_{0}(1-N / K)
$$
$r_0$ is called maximal per capita reproduction rate.\\
Then the dynamic equation is
\begin{align}
    \mathrm{d} N / \mathrm{d} t=r_{0} N(1-N / K) \label{eq1.7}
\end{align}
\myref{eq1.7} is non-linear.
\newpage
\paragraph{Solution}~{}\\
Solve \myref{eq1.7} with initial value $N(0)=N_0$.
\subparagraph{When $N_0=0$, $N(t)=0$ for $t \in (0,\infty) $}
\subparagraph{When $N_0>0, N(t) \rightarrow K $ as $ t \rightarrow+\infty$.}
\begin{gather*}
    \mathrm{d} N / \mathrm{d} t=r N(1-N / K)\\
    N(t)=\frac{K}{1+\left(K / N_{0}-1\right) e^{-r t}}
\end{gather*}
Graph!\\
K is called \emph{\textbf{carrying capacity}}.

\newpage
\subsubsection{Richards Model}
\paragraph{Equation}~{}\\
Another guess of $r$ is 
$$
r=r_0\left(1-(N / K)^{\nu}\right)
$$
$$
v =\text{ constant } >0
$$
Then the dynamic equation is
\begin{align}
    \mathrm{d} N / \mathrm{d} t=r N(1-(N / K)^v) \label{eq1.8}
\end{align}

\paragraph{Solution}~{}\\
Solve \myref{eq1.8} using substitution (i.e. let $u=(\frac{N}{K})^{-v})$. 

\begin{equation*}
    N(t)= \begin{cases}\frac{k}{\left[\left(\left(\frac{k}{N_{0}}\right)^{\nu}-1\right) e^{-v r t}+1\right]^{\frac{1}{v}}} &,N(0)\ne 0 \text{ or } k \\ 0&,N(0)=0 \\ k&,N(0)=k\end{cases}
\end{equation*}


\newpage

\subsubsection{General case}
Any \emph{\textbf{single-species continuous time}} equation is solved in \emph{\textbf{quadratures}}
\begin{align*}
    \mathrm{d} N / \mathrm{d} t=r(N) N=f(N)\\
    \int_{N_{0}}^{N} \frac{\mathrm{d} N}{f(N)}=\int_{0}^{t} \mathrm{~d} t=t
\end{align*}
with special solutions (equilibria).

\subsection{Single Species continuous Model: qualitative analysis}


\newpage

\subsection{Spruce Budworm outbreak}